A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration.

### 2024-01-31 frontiers_submission/frontiers_submission.Rmd
finish regressions

---
title: "A meta-analysis of the impacts of best management practices on nonpoint source pollutants."
runtitle: "Meta-analysis of best management practices"
documentclass: FrontiersinHarvard # for articles in journals using the Harvard Referencing Style (Author-Date), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: FrontiersinVancouver # for articles in journals using the Vancouver Reference Style (Numbered), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: frontiersinFPHY_FAMS # Vancouver Reference Style (Numbered) for articles in the journals "Frontiers in Physics" and "Frontiers in Applied Mathematics and Statistics" 
classoption: utf8
author:
  - name: Michael Schramm
    affiliation: '1'
    etal: Schramm # First author's last name. 
    email: michael.schramm@ag.tamu.edu # Indicates corresponding Author
    institution: Texas Water Resources Institute, Texas A&M AgriLife Research
    street: 1001 Holleman Dr E.
    city: College Station
    state: Texas # only USA, Australia, Canada
    zip: 77840
    country: USA
  - name: Duncan Kikoyo
    affiliation: '1'
  - name: Janelle Wright
    affiliation: '1'
  - name: Shubham Jain
    affiliation: '1'
affiliation:
  - id: '1'
    department: Texas Water Resources Institute
    institution: Texas A&M AgriLife Research
    city: College Station
    state: Texas # only USA, Australia, Canada
    country: USA

# Abstract can also be provided at the 
abstract: |
  Abstract length and content varies depending on article type. Refer to 
  <http://www.frontiersin.org/about/AuthorGuidelines> for abstract requirement
  and length according to article type.
keywords:
  - best management practice
  - water quality
  - nonpoint source pollution
  - fecal indicator bacteria
  - nutrients
output: rticles::frontiers_article
# bib yaml is commented to manually insert in correct section
# csl: frontiers.csl
# bibliography: bibliography.bib
# Match bibliostyle with the documentclass
# biblio-style: Frontiers-Harvard # Many Frontiers journals use the Harvard referencing system (Author-date), to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal. For Humanities and Social Sciences articles please include page numbers in the in-text citations 
# biblio-style: Frontiers-Vancouver # Many Frontiers journals use the numbered referencing system, to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
---

```{r setup, include=FALSE}
## load libraries
library(flextable)
library(ggtext)
library(kableExtra)
library(officer)
library(ragg)
library(targets)
library(tidyverse)

@@

## grab output from targets to use in manuscript
store <- "../_targets"
bac_missing_sd <- tar_read(bac_missing_sd,
                           store = store)
tn_missing_sd <- tar_read(tn_missing_sd,
                           store = store)
tp_missing_sd <- tar_read(tp_missing_sd,
                           store = store)
tss_missing_sd <- tar_read(tss_missing_sd,
                           store = store)

eggers_bac <- tar_read(bac_sse,
                       store = store)
## set flextable defaults
set_flextable_defaults(fonts_ignore=TRUE)

@@

In the United States (U.S.), major progress towards water quality goals have been achieved through the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@benhamLessonsLearnedTMDL2008; @nationalresearchcouncilAssessingTMDLApproach2001; @schrammTotalMaximumDaily2022].
Increased pollutant loads and pollutant concentrations in runoff are a particular challenge when dealing with nonpoint source pollution resulting from land use change.
The impacts of land use change on hydrology and water quality are well established [@allanLandscapesRiverscapesInfluence2004; @carpenterNonpointPollutionSurface1998; @bernhardtUnderstandingManagingMinimizing2008; @careyEvaluatingNutrientImpacts2013; @freemanImpactsUrbanizationDevelopment2019].
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus [@epaNationalWaterQuality2017]. 
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States [@epaNationalWaterQuality2017].

@@

Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
@linternBestManagementPractices2020 found 43% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation. 
However, field studies are much more likely to demonstrate mixed results including net releases of pollutants under certain conditions  [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of parameters that represent management practices [@ullrichApplicationSoilWater2009; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], (2) failure to incorporate uncertainty in estimates [@tasdighiBayesianTotalUncertainty2018; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], and (3) assumption of static performance over time [@mealsLagTimeWater2010; @liuReviewEffectivenessBest2017; @fuReviewCatchmentscaleWater2019].

A source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies [@linternBestManagementPractices2020].
Previous empirical reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types [@linternBestManagementPractices2020;@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014; @claryBMPPerformanceAnalysis2011; @barrettComparisonBMPPerformance2008; @grudzinskiDoesRiparianFencing2020]. 
The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs [@claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020].
Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
However, influent water quality, seasonality, and BMP age provide some explanatation to highliy heterogenous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].  
A recent meta-analysis demonstrated the further linkages between local climatic conditions and BMP performance on nitrogen and phosphorus concentrations [@horvathEffectsRegionalClimate2023].

@@

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations. 
The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “E. coli” OR “Escherichia coli” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”. Nutrient BMP studies utilized a similar query: “nutrient” OR “nitrogen” OR “phosphorus” OR “sediment” OR “TSS” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included.
**include supplementary tables with inclusion criteria and data variables**
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction, again with at least two team members reviewing each study. 
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022].

@@

tibble(Variable = c(
  "Publication Year", "Parameter", "Runoff source", "Source type",
  "BMP", "BMP Classification", "BMP Category", "BMP Subcategory",
  "Study scale", "Location", "State", "Study area", "Longitude", "Latitude",
  "Study years", "N control", "N experiment", "X control", "X experiment", 
  "SE control", "SE experiment", "Minimum control", "Minimum experiment", 
  "Maximum control", "Maximum experiment", "SD control", "SD experiment",
  "Units", "Percent reduction"),
  Description = c("Year the study was published",
                  "The specific pollutant measured.",
                  "Dominant source of runoff (rop fields, livestock pasture, commerical, residential",
                  "Major categorization of runoff source: agricultural or urban",
                  "BMP evaluated",
                  "BMP description based on NRCS conservation practice standards and EPA BMP fact sheets",
                  "BMP categorization based on structural or management",
                  "BMP subcategorization based on pollutant removal processes",
                  'Spatial scale of the study area ("lot/field", "community", "watershed")',
                  "Location name used in the study description",
                  "State where the study was conducted",
                  "Drainage area in hectares",
                  "Approximated or reported latitude coordinate",
                  "Approximated or reported longitude coordinate",
                  "Year or years when data were collected",
                  "Number of control measurements",
                  "Number of experiemental measurements",
                  "Mean concentration for control measurements",
                  "Mean concentration for experiemental measurements",
                  "Standard error of control measurements",
                  "Standard error of experimental measurements",
                  "Minimum control measurement",
                  "Minimum experiment measurement",
                  "Maximum control measurement",
                  "Maximum experiment measurement",
                  "Standard deviation of control measreuments",
                  "Standard deviation of experimental measurements",
                  "Units reported by the study",
                  "BMP efficiency for studies that only reported efficiency")
) |> 
  kableExtra::kbl(booktabs = TRUE,
                  caption = "Study and effect variables extracted for review.") |> 
  kableExtra::kable_styling(full_width = FALSE) |>
  kableExtra::column_spec(1, width = "10em") |> 
  kableExtra::column_spec(2, width = "25em")

@@

## Effect size calculation {-}

Prior BMP reviews routinely report the effect of BMPs using efficiency values [see @agouridisLivestockGrazingManagement2005; @krogerReviewBestManagement2012; @liuEnhancingRainfallrunoffModel2015; @simpsonDevelopingBestManagement2009].
Efficiency values are calculated as:

$$
\text{Efficiency}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
$$

where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
There are several statistical shortcomings (distributional asymetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
In comparision, the log ratio of means (*ROM*) provide preferable statistical properties for regression analysis [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
*ROM* quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:

$$
ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
$$
where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
The statistical properties of *lnRR* (normal distribution around zero and additive properties) are preferable to the more commonly reported efficiency metric [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999]. 
Use of *ROM* required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.

@@

We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects linear models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
`r paste0(round(bac_missing_sd$missing / bac_missing_sd$total * 100, 0), "%")` of fecal indicator bacteria, `r paste0(round(tn_missing_sd$missing / tn_missing_sd$total * 100, 0), "%")` of TN, `r paste0(round(tp_missing_sd$missing / tp_missing_sd$total * 100, 0), "%")` of TP, and `r paste0(round(tss_missing_sd$missing / tss_missing_sd$total * 100, 0), "%")` of TSS effect sizes were missing standard deviations required to estimate sampling variance. 
Removal of studies due to missing variance information can introduce substantial bias [@kambachConsequencesMultipleImputation2020], so we imputed missing standard deviations using random forest based multivariate imputation with chained equations using the *mice* R package [@buurenMiceMultivariateImputation2011].
Sampling variance was estimated with the variance estimator described in Doncaster et al. to reduce bias in small sample studies:

@@

Our initial models included log transformed pre-treatment pollutant concentration, study scale (plot/field, community, watershed), study length (*n* years), BMP subcategory (), and pollutant concentration:BMP subcategory interactions were included as fixed effect terms.
We used an information-theoretic approach to select the most parsimonious model from the subset of candidate models based on corrected Akaike information criterion (AIC~c~) estimated with maximum likelihood [@cinarUsingInformationTheoretic2021].
The final model was selected from candidate models, which included all combination and subsets of the full model, by eliminating candidate models until the $\Delta$*AIC~c~* reached $\leq$ 2 [@burnhamAICModelSelection2011].
We estimated regression coefficients of the selected model using restricted maximum likelihood.
Relative heterogeneity between and within studies were calculated using the *I^2^* metric described in @nakagawaMethodologicalIssuesAdvances2012.
Marginal *R^2^* was used to describe the amount of variance explained by fixed effects [@nakagawaGeneralSimpleMethod2013].

```{r}
bac_z <- eggers_bac$zval[[3]] |> prettyNum(digits = 3, nsmall = 2)
bac_p <- eggers_bac$pval[[3]] |> format.pval(eps = .01, digits = 2, nsmall = 1)
```

We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023]. 
We did not identify evidence of publication bias in the surveyed studies for fecal indicator bacteria (*Z* = `r bac_z`, *p* = `r bac_p`), total nitrogen, total phosphorus, or total suspended solids; therefore, adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
Outlier studies identified in the sensitivity analysis are highlighted in Table X and were removed prior to fitting the final models.

>>>>

---
title: "A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration."
runtitle: "Meta-analysis of best management practices"
documentclass: FrontiersinHarvard # for articles in journals using the Harvard Referencing Style (Author-Date), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: FrontiersinVancouver # for articles in journals using the Vancouver Reference Style (Numbered), for Frontiers Reference Styles by Journal: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
# documentclass: frontiersinFPHY_FAMS # Vancouver Reference Style (Numbered) for articles in the journals "Frontiers in Physics" and "Frontiers in Applied Mathematics and Statistics" 
classoption: utf8
author:
  - name: Michael Schramm
    affiliation: '1'
    etal: Schramm # First author's last name. 
    email: michael.schramm@ag.tamu.edu # Indicates corresponding Author
    institution: Texas Water Resources Institute, Texas A&M AgriLife Research
    street: 1001 Holleman Dr E.
    city: College Station
    state: Texas # only USA, Australia, Canada
    zip: 77840
    country: USA
  - name: Duncan Kikoyo
    affiliation: '1'
  - name: Janelle Wright
    affiliation: '1'
  - name: Shubham Jain
    affiliation: '1'
affiliation:
  - id: '1'
    department: Texas Water Resources Institute
    institution: Texas A&M AgriLife Research
    city: College Station
    state: Texas # only USA, Australia, Canada
    country: USA

# Abstract can also be provided at the 
abstract: |
  Abstract length and content varies depending on article type. Refer to 
  <http://www.frontiersin.org/about/AuthorGuidelines> for abstract requirement
  and length according to article type.
keywords:
  - best management practice
  - water quality
  - nonpoint source pollution
  - fecal indicator bacteria
  - nutrients
output: 
  bookdown::pdf_book:
    base_format: rticles::frontiers_article
    pandoc_args:  !expr rmdfiltr::add_wordcount_filter()
# bib yaml is commented to manually insert in correct section
# csl: frontiers.csl
# bibliography: bibliography.bib
# Match bibliostyle with the documentclass
# biblio-style: Frontiers-Harvard # Many Frontiers journals use the Harvard referencing system (Author-date), to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal. For Humanities and Social Sciences articles please include page numbers in the in-text citations 
# biblio-style: Frontiers-Vancouver # Many Frontiers journals use the numbered referencing system, to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
---

```{r setup, include=FALSE}
## load libraries
library(ggtext)
library(kableExtra)
library(officer)
library(ragg)
library(targets)
library(tidyverse)

@@

## grab output from targets to use in manuscript
store <- "../_targets"
bac_missing_sd <- tar_read(bac_missing_sd,
                           store = store)
tn_missing_sd <- tar_read(tn_missing_sd,
                           store = store)
din_missing_sd <- tar_read(tin_missing_sd,
                           store = store)
tp_missing_sd <- tar_read(tp_missing_sd,
                           store = store)
po4_missing_sd <- tar_read(po4_missing_sd,
                           store = store)
tss_missing_sd <- tar_read(tss_missing_sd,
                           store = store)

missing_sd <- list(bac_missing_sd,
                tn_missing_sd,
                din_missing_sd,
                tp_missing_sd,
                po4_missing_sd,
                tss_missing_sd) 
missing <- missing_sd |>
  map_dbl("missing") |> 
  sum()
total <- missing_sd |> 
  map_dbl("total") |> 
  sum()

percent_missing <- round(missing/total * 100,0)

fib_r2 <- tar_read(bac_r2, store = store)
fib_i2 <- tar_read(bac_i2, store = store)
fib_i2_null <- tar_read(bac_i2_null, store = store)

tn_i2_null <- tar_read(tn_i2_null, store = store)
din_i2_null <- tar_read(tin_i2_null, store = store)

tp_i2_null <- tar_read(tp_i2_null, store = store)
po4_i2_null <- tar_read(po4_i2_null, store = store)

tss_i2_null <- tar_read(tss_i2_null, store = store)

@@

In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge [@benhamLessonsLearnedTMDL2008; @nationalresearchcouncilAssessingTMDLApproach2001; @schrammTotalMaximumDaily2022].
Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
The impacts of land use change on hydrology and water quality are well established [@allanLandscapesRiverscapesInfluence2004; @carpenterNonpointPollutionSurface1998; @bernhardtUnderstandingManagingMinimizing2008; @careyEvaluatingNutrientImpacts2013; @freemanImpactsUrbanizationDevelopment2019].
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus [@epaNationalWaterQuality2017]. 
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States [@epaNationalWaterQuality2017].

@@

Practitioners rely extensively on mechanistic models to plan and evaluate BMP scenarios and resulting water quality.
@linternBestManagementPractices2020 found 43% of reviewed BMP effectiveness studies relied completely on modeled outputs, with modeled outputs almost always predicting water quality improvements following BMP implementation. 
However, field studies are much more likely to demonstrate mixed results including net releases (leaching) of pollutants under certain conditions  [@linternBestManagementPractices2020; @liuReviewEffectivenessBest2017].
The disconnect between modeled outcomes and field studies might be attributed to (1) overly simplified or incorrect estimates of model parameters that represent management practices [@ullrichApplicationSoilWater2009; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], (2) the failure to incorporate the appropriate types of uncertainty into estimates [@tasdighiBayesianTotalUncertainty2018; @fuReviewCatchmentscaleWater2019; @linternBestManagementPractices2020], and (3) the assumption of static performance over time [@mealsLagTimeWater2010; @liuReviewEffectivenessBest2017; @fuReviewCatchmentscaleWater2019].

An underlying source of uncertainty comes from the substantial variability of performance metrics reported in empirical BMP studies [@linternBestManagementPractices2020].
There have been varied attempts at synthesizing estimates of BMP efficiency to provide resource managers with knowledge for improved decision-making [@agouridisLivestockGrazingManagement2005; @barrettComparisonBMPPerformance2008; @claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020; @horvathEffectsRegionalClimate2023; @kochNitrogenRemovalStormwater2014; @krogerReviewBestManagement2012; @liuReviewEffectivenessBest2017; @simpsonDevelopingBestManagement2009]. 
These reviews generally describe high variability and uncertainty in nitrogen and phosphorus removal and consistent reduction in total suspended sediment concentrations across BMP types [@linternBestManagementPractices2020;@liuReviewEffectivenessBest2017; @kochNitrogenRemovalStormwater2014; @claryBMPPerformanceAnalysis2011; @barrettComparisonBMPPerformance2008; @grudzinskiDoesRiparianFencing2020]. 
The review literature on the effects of BMPs on FIBs are sparse but generally find extremely high variance in performance across BMPs [@claryBMPPerformanceAnalysis2011; @grudzinskiDoesRiparianFencing2020].

## Sources of variance in BMP performance

While it is assumed that site specific conditions are responsible for some of the heterogeneity in observed BMP performance, it is not clear how much of that variance is due to any one specific factor.
Influent concentration is likely to have some effect on certain types of structural BMPs.
@barrettPerformanceComparisonStructural2005 demonstrated that percent pollutant reduction is often a function of influent quality.
Specifically, for certain types of BMPs percent removal is low at low influent concentrations, and increases with increasing influent concentrations. 
However, for some types of BMPs and pollutant parameters, effluent concentration is unrelated to influent concentration.
Second, local climatic conditions can be expected to influence BMP performance.
BMPs in dry climates have been shown to be more likely to leach phosphorus than those in wetter climates [@horvathEffectsRegionalClimate2023].
However, elucidating possible confounders such as climate and soil condition has been constrained by the lack of reported local condition data included in most BMP studies [@horvathEffectsRegionalClimate2023; @kochNitrogenRemovalStormwater2014; @eagleMetaAnalysisConstrained2017].
A third variable that has been explored is the infuence of BMP age and maintence on performance.

Accurately characterizing BMP treated runoff quality is challenging due to the high variability in BMP performance.
However, influent water quality, seasonality, and BMP age provide some explanation to highly heterogeneous BMP efficiency data [@barrettPerformanceComparisonStructural2005; @liuReviewEffectivenessBest2017].  
A recent meta-analysis demonstrated the strong linkages between local climatic conditions and phosphorus removal in BMPs [@horvathEffectsRegionalClimate2023].

## Effect size calculations

Results from BMP studies are often reported as BMP efficiency (or percent reduction):

$$
\text{BMP}_{\text{eff}}=\frac{x_{control}-x_{experiment}}{x_{control}}\times 100,
$$

where *x~control~* is the pre-treatment or control pollutant concentration and *x~experiment~* is the pollutant concentration measured after the BMP intervention. 
Several BMP data synthesis efforts have applied statistical summaries or regressions using BMP efficiency as the response variable of interest [@agouridisLivestockGrazingManagement2005; @claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @krogerReviewBestManagement2012; @liuReviewEffectivenessBest2017; @simpsonDevelopingBestManagement2009].
There are several statistical shortcomings (distributional asymmetry, skewness, and non-additive properties) when using efficiency to estimate overall effect sizes across multiple studies that are cause for concern for metrics estimated using this approach [@nuzzoPercentDifferencesAnother2018; @coleStatisticsNotesWhat2017].
@barrettPerformanceComparisonStructural2005 demonstrated the use of effluent concentration directly as a response variable improved the ability to describe BMP performance. 
More recently, researchers have applied effect size calculations more commonly used in ecological meta-analysis.
@horvathEffectsRegionalClimate2023 used the standardized mean difference between influent and effluent, calculated as the difference in the means divided by the pooled standard deviation of the two groups (**hedges**).
@grudzinskiDoesRiparianFencing2020 applied the log ratio of means (*ROM*) to summarize performance of livestock BMPs.
*ROM* quantifies the difference in means between the control and experimental group [@hedgesMetaanalysisResponseRatios1999]:

$$
ROM_i = ln\left(\frac{x_{i,control}}{x_{i,experiment}}\right) = ln(x_{i, control})-ln(x_{i, experiment}),
$$

where *x~i,control~* and *x~i,experiment~* are the mean pollutant concentrations for experiment *i*. 
The statistical properties of *ROM* (normal distribution around zero and additive properties) are preferable to using BMP efficiency [@osenbergEffectSizeEcological1997; @hedgesMetaanalysisResponseRatios1999].
*ROM* \textgreater 0 indicates higher percent reductions and *ROM* \textless indicates pollutant leaching.
One advantage of *ROM* is that the statistical results calculated using *ROM* are easily transformed to BMP efficiency for interpretation:

$$
\text{BMP}_{\text{eff}} = \left( 1 - \frac{1}{e^{ROM}}\right) \times 100
$$

**still needs work**
Building on previous work, the objective of this project is to (1) improve understanding of the general effectiveness of BMPs on pollutant concentration, and (2) identify important contributors to variability in BMP performance. 
To accomplish this, we conducted a systematic review of relevant published literature and applied a meta-analytic approaches to develop weighted results across studies and identify contributors to variance in BMP performance.

@@

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations. 
Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014; @horvathEffectsRegionalClimate2023].
The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data [@claryBMPPerformanceAnalysis2011; @kochNitrogenRemovalStormwater2014].
Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.

The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines [@collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018].
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: “fecal indicator bacteria” OR “*E. coli*” OR “*Escherichia coli*” OR “enterococci” OR “enterococcus” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”. Nutrient BMP studies utilized a similar query: “nutrient” OR “nitrogen” OR “phosphorus” OR “sediment” OR “TSS” AND “best management practices” OR “BMPs” AND “effectiveness” OR “performance”.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (*n* = 4) was assigned a subset of studies to evaluate if they should be included (Table \@ref(tab:criteria)).
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction (Table \@ref(tab:datvars)), again with at least two team members reviewing each study. 
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool [@rohatgiWebPlotDigitizer2022].

@@

tibble(Variable = c(
  "Publication Year", "Parameter", "Runoff source", "Source type",
  "BMP", "BMP Classification", "BMP Category", "BMP Subcategory",
  "Study scale", "Location", "State", "Study area", "Longitude", "Latitude",
  "Study years", "N control", "N experiment", "X control", "X experiment", 
  "SE control", "SE experiment", "Minimum control", "Minimum experiment", 
  "Maximum control", "Maximum experiment", "SD control", "SD experiment",
  "Units", "Percent reduction"),
  Description = c("Year the study was published",
                  "The specific pollutant measured.",
                  "Dominant source of runoff (crop fields, livestock pasture, commerical, residential",
                  "Major categorization of runoff source: agricultural or urban",
                  "BMP evaluated",
                  "BMP description based on NRCS conservation practice standards and EPA BMP fact sheets",
                  "BMP categorization based on structural or management",
                  "BMP subcategorization based on pollutant removal processes",
                  'Spatial scale of the study area ("lot/field", "community", "watershed")',
                  "Location name used in the study description",
                  "State where the study was conducted",
                  "Drainage area in hectares",
                  "Approximated or reported latitude coordinate",
                  "Approximated or reported longitude coordinate",
                  "Year or years when data were collected",
                  "Number of control measurements",
                  "Number of experiemental measurements",
                  "Mean concentration for control measurements",
                  "Mean concentration for experiemental measurements",
                  "Standard error of control measurements",
                  "Standard error of experimental measurements",
                  "Minimum control measurement",
                  "Minimum experiment measurement",
                  "Maximum control measurement",
                  "Maximum experiment measurement",
                  "Standard deviation of control measreuments",
                  "Standard deviation of experimental measurements",
                  "Units reported by the study",
                  "BMP efficiency for studies that only reported efficiency")
) |> 
  kableExtra::kbl(booktabs = TRUE,
                  caption = "Study and effect variables extracted for review.") |> 
  kableExtra::kable_styling(full_width = FALSE) |>
  kableExtra::column_spec(1, width = "10em") |> 
  kableExtra::column_spec(2, width = "25em")

@@

@@

We used the "rma.mv" function in the *metafor* R package to fit multilevel random effects regression models with *ROM* as the effect variable [@viechtbauerConductingMetaanalysesMetafor2010; @rcoreteamLanguageEnvironmentStatistical2023].
Our models specified a nested random effects term accounting for heterogeneity between effect sizes from the same study and for heterogeneity between studies.
*ROM* was used as the effect size which required the exclusion of studies that only provided measures of BMP efficiency and not the underlying data used to derive the metric.
A key feature of meta-analysis is the weighting of effects using sampling variance of individual effect sizes.
`r percent_missing |> xfun::numbers_to_words(cap = TRUE)` percent of `r total` effect sizes were missing standard deviations required to estimate sampling variance. 
Removal of studies due to missing variance information can introduce substantial bias [@kambachConsequencesMultipleImputation2020]. 
Missing standard deviations were imputed using the pooled ratio of the mean effect size to coefficient of variation (CV) [@brackenStatisticalMethodsAnalysis1992].
Sampling variance was estimated utilizing the average squared CV across all studies divided by sample size for each effect [@nakagawaRobustReadilyImplementable2023; @doncasterCorrectionBiasMeta2018]:

@@

Our initial models included log transformed influent concentration, BMP subcategory (drainage modification, crop field management, livestock management, filtration, treatment, detention, or infiltration), aridity index (mean-centered), influent concentration\texttimes BMP subcategory interactions, and aridity index\texttimes BMP subcategory interactions were included as fixed effect terms.
Aridity index was the only moderator not obtained directly in the systematic review (Table \@ref(tab:datvars)).
We mapped study location coordinates to aridity index values published in the "Global Aridity Index and Potential Evapotranspiration Database - Version 3" (Global-AI_PET_v3) which provides gridded 30 arc-second annual average precipitation and potential evapotranspiration estimates [@zomerVersionGlobalAridity2022].
The aridity index is calculated as the ratio of mean annual precipitation to mean annual evapotraspiration with values between 0 and 0.5 considered hyper to semi-arid, and values above 0.65 as humid.

We used an information-theoretic approach to select the most parsimonious model from the subset of candidate models based on corrected Akaike information criterion (AIC~c~) estimated with maximum likelihood [@cinarUsingInformationTheoretic2021].
Candidate models used for variable selection were fit with maximum likelihood (ML).
The final model was selected from candidate models, which included all combination and subsets of the full model, by selecting the model with the lowest AIC~c~ score [@burnhamAICModelSelection2011; @cinarUsingInformationTheoretic2021].
Regression coefficients of the selected model were estimated using restricted maximum likelihood (REML).
Relative heterogeneity between and within studies were calculated using the *I^2^* metric described in @nakagawaMethodologicalIssuesAdvances2012.
Marginal *R^2^* was used to describe the amount of variance explained by fixed effects [@nakagawaGeneralSimpleMethod2013].

We tested for evidence of publication bias, in the form of small study effect, by using the extension of Egger's regression applied to the multilevel model framework that included adjusted sampling error as a moderator [@nakagawaQuantitativeEvidenceSynthesis2023]. 
We did not identify evidence of publication bias in the surveyed studies (FIB: ROM = 1.19, 95% CI [-3.23, 5.61]; TN: ROM = 0.31, 95% CI [-1.15, 1.78]; DIN: ROM = -2.4, 95% CI [-6.14, 1.59]; TP: ROM = -1.97, 95% CI [-5.18, 1.23]; PO~4~: -0.05, 95% CI [-2.82, 2.72]; TSS: ROM: 0.31, 95% CI [-1.15, 1.78]; Figure S1-S6); therefore, adjustments for publication bias were not included in the final models.
We conducted a sensitivity analysis of the robustness of overall effect sizes to individual studies using leave-one-out analysis [@nakagawaQuantitativeEvidenceSynthesis2023].
This approach repeatedly fits the selected model leaving out an individual value each time. 
The overall effect and 95% CI from each refit model is compared to the overall effect and 95% CI of the model fit to the full dataset.
We did not identify evidence of outliers or overly influential studies for any of our models (Figure S7-S12).
